\documentclass{article}
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage[margin=0.5in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts, fancyhdr, color, comment, graphicx, environ}
\usepackage{float}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage[shortlabels]{enumitem}
\usepackage{indentfirst}
\usepackage{mathrsfs}
\usepackage{hyperref}
\graphicspath{./}
\makeatletter
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother
% Change enumerate labels to (a), (b), (c), ...
% Define a new environment for problems
\newcounter{problemCounter}
\newtcolorbox{problem}[2][]{colback=white, colframe=black, boxrule=0.5mm, arc=4mm, auto outer arc, title={\ifstrempty{#1}{Problem \stepcounter{problemCounter}\theproblemCounter}{#1}}}

% \renewcommand{\labelenumi}{\alph{enumi})}
\def\zz{{\mathbb Z}}
\def\rr{{\mathbb R}}
\def\qq{{\mathbb Q}}
\def\cc{{\mathbb C}}
\def\nn{{\mathbb N}}
\def\ss{{\mathbb S}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtcolorbox{proposition}[1][]{colback=white, colframe=blue, boxrule=0.5mm, arc=4mm, auto outer arc, title={Proposition #1}}
\newtcolorbox{definition}[1][]{colback=white, colframe=violet, boxrule=0.5mm, arc=4mm, auto outer arc, title={Definition #1}}
\newcommand{\Zmod}[1]{\zz/#1\zz}
\newcommand{\partFrac}[2]{\frac{\partial #1}{\partial #2}}

\newcommand\Mydiv[2]{%
$\strut#1$\kern.25em\smash{\raise.3ex\hbox{$\big)$}}$\mkern-8mu
        \overline{\enspace\strut#2}$}

\begin{document}

\begin{center}
    Math 540
    \hfill Homework 1
    \hfill \textit{Stephen Cornelius}
\end{center}
\textbf{Remarks:} \\
\begin{enumerate}[A)]
    \item Definition is just a definition, there is no need to jjustify or explain it.
    \item Answers to questions with proofs should be written, as much as you can, in the following format: \\
    \begin{enumerate}[i)]
        \item Statement
        \item Main points that will appear in your proof
        \item The actual proof
    \end{enumerate}
    Answers to questions with computations should be written, as much as possible, in the following format:
    \begin{enumerate}[i)]
        \item Statement and Result
        \item Main points that will appear in your computation.
        \item The actual computation
    \end{enumerate}
\end{enumerate}



% Start of problems




\begin{problem}
    \textit{Vector Spaces}. Suppose $\mathbb{F}$ is a field.
    \begin{enumerate}[a)]
        \item Define when we say that a vector space $V$ over a field $\mathbb{F}$ is \textit{finite dimensional}.
        \item Consider the vector space
        \[
            V = \mathbb{F}[x]
        \]
        of all polynomials with coefficients in $\mathbb{F}$. Show that $V$ is not finite dimensional.
        \item Suppose $X$ is a finite set. Consider the vector space $V$, of all functions from $X$ to $\mathbb{F}$,
        \[
            V = \mathbb{F}(X):= \{\text{all }f: X \to \mathbb{F}; \text{s.t. $f$ is a function}\},
        \]
        with the standard addition and multiplication by scalars from $\mathbb{F}$. Show that $V$ is finite dimensional.
    \end{enumerate}
\end{problem}

\begin{enumerate}[a)]
    \item We say that a vector space $V$ over a field $\mathbb{F}$ is finite dimensional if there exists $S \subseteq V$ such that $\#(S) < \infty$ and $\operatorname{span}(S) = V$.
    \item \begin{enumerate}[i)]
        \item Statement: Show that $V = \mathbb{F}[x]$ is not finite dimensional.
        \item Main Points:
        \begin{itemize}
            \item Suppose $V$ is finite dimensional with a finite spanning set $S$.
            \item Let $m$ be the maximum degree of the polynomials in $S$ and find a polynomial $q(x)$ with degree greater than $m$.
            \item Show that $q(x)$ cannot be written as a linear combination of the polynomials in $S$, leading to a contradiction.
        \end{itemize}
        \item Proof: 
        \begin{proof}
            Suppose $V$ is finite dimensional. Then there exists $S \subseteq V$ such that $\#(S) < \infty$ and $\operatorname{span}(S) = V$. Let $S = \{p_1(x), p_2(x), \ldots, p_n(x)\}$. Let $m = \max(\deg(p_i(x)))$ for $1 \leq i \leq n$. Then consider the polynomial $q(x) = x^{m+1}$. Since $\deg(q(x)) > m$, $q(x)$ cannot be written as a linear combination of the polynomials in $S$. This contradicts the fact that $\operatorname{span}(S) = V$. Thus, $V$ is not finite dimensional.
        \end{proof}
        \end{enumerate}

        \item \begin{enumerate}[i)]
            \item Statement: Show that $V = \mathbb{F}(X)$ is finite dimensional.
            \item Main Points:
                \begin{itemize}
                    \item Since $X$ is finite, say $X = \{x_1, x_2, \ldots, x_n\}$, consider the set of functions $\{\delta_{x_i}\}_{i=1}^n$, where
                    \[
                        \delta_{x_i}(x_j) = \begin{cases}
                            1 & \text{if } i = j \\
                            0 & \text{if } i \neq j
                        \end{cases}
                    \]
                    \item Show that $\{\delta_{x_i}\}_{i=1}^n$ forms a basis for $V$.
                \end{itemize}
            \item Proof:
            \begin{proof}
                Since $X$ is finite, say $X = \{x_1, x_2, \ldots, x_n\}$, consider the set of functions $\{\delta_{x_i}\}_{i=1}^n$, where
                \[
                    \delta_{x_i}(x_j) = \begin{cases}
                        1 & \text{if } i = j \\
                        0 & \text{if } i \neq j
                    \end{cases}
                \]
                We will show that $\{\delta_{x_i}\}_{i=1}^n$ forms a basis for $V$. First, we show that they span $V$. Let $f \in V$. Then we can write
                \[
                    f(x) = \sum_{i=1}^n f(x_i) \delta_{x_i}(x)
                \]
                for all $x \in X$. Thus, $\{\delta_{x_i}\}_{i=1}^n$ spans $V$. Next, we show that they are linearly independent. Suppose
                \[
                    \sum_{i=1}^n a_i \delta_{x_i}(x) = 0
                \]
                for some scalars $a_i \in \mathbb{F}$. Evaluating at $x = x_j$, we get
                \[
                    a_j = 0
                \]
                for all $j = 1, 2, \ldots, n$. Thus, all $a_i = 0$, and $\{\delta_{x_i}\}_{i=1}^n$ are linearly independent. Therefore, $\{\delta_{x_i}\}_{i=1}^n$ forms a basis for $V$, and hence $V$ is finite dimensional with $\dim(V) = n$.
            \end{proof}
        \end{enumerate}
\end{enumerate}





\begin{problem}
    \textit{Short exact sequences}. Suppose $U,V,W$ are three vector spaces over $\mathbb{F}$. Consider the following sequence of spaces and linear transformations between them:
    \begin{equation}
        0 \xrightarrow{} U \xrightarrow{\iota} V \xrightarrow{\epsilon} W \xrightarrow{} 0,
    \end{equation}
    where $0 \to U$, are the obvious maps from the zero space into $U$, and from the space $W$ onto the zero space, respectively. \\
    \begin{enumerate}[a)]
        \item Define when we say that the sequence (1) is \underline{short exact sequence} (s.e.s.).
        \item Given two subspaces $U,V < V$, such that $V = U \oplus W$, Show that there is a natural s.e.s.associated with the spaces of functions $U = \mathbb{F}(U), V = \mathbb{F}(V)$ and $W = \mathbb{F}(Y \backslash X)$, where $Y \backslash X$ denotes set-minus, i.e., the set of elements which are in $Y$ and are not in $X$.
    \end{enumerate}
\end{problem}


\begin{enumerate}[a)]
    \item A sequence (1) is a short exact sequence (s.e.s.) if the image of each map is equal to the kernel of the next map, i.e.,
    \[
        \operatorname{im}(0 \to U) = \ker(\iota), \quad \operatorname{im}(\iota) = \ker(\epsilon), \quad \operatorname{im}(\epsilon) = \ker(0 \to W).
    \]
    Since the maps from and to the zero space are trivial, this reduces to
    \[
        \iota \text{ is injective}, \quad \epsilon \text{ is surjective}, \quad \operatorname{im}(\iota) = \ker(\epsilon).
    \]
    \item \begin{enumerate}[i)]
        \item Statement: There is a natural s.e.s. associated with the spaces of functions $U = \mathbb{F}(U), V = \mathbb{F}(V)$ and $W = \mathbb{F}(Y \backslash X)$.
        \item Main Points:
            \begin{itemize}
                \item Define the inclusion map $\iota: U \to V$ by extending functions by zero outside $U$.
                \item Define the projection map $\epsilon: V \to W$ by restricting functions to $W$.
                \item Show that $\iota$ is injective.
                \item Show that $\epsilon$ is surjective.
                \item Show that $\operatorname{im}(\iota) = \ker(\epsilon)$.
            \end{itemize}
        \item Proof:
        \begin{proof}
            Since $V = U \oplus W$, every element $v \in V$ can be uniquely written as $v = u + w$ for some $u \in U$ and $w \in W$. Define the inclusion map $\iota: U \to V$ by
            \[
                \iota(f)(x) = \begin{cases}
                    f(x) & \text{if } x \in U \\
                    0 & \text{if } x \notin U
                \end{cases}
            \]
            for all $f \in U$. This map is injective because if $\iota(f) = 0$, then $f(x) = 0$ for all $x \in U$, which implies $f = 0$. Next, define the projection map $\epsilon: V \to W$ by
            \[
                \epsilon(g)(x) = g(x)
            \]
            for all $g \in V$ and $x \in W$. This map is surjective because for any $h \in W$, we can define a function $g \in V$ by
            \[
                g(x) = \begin{cases}
                    h(x) & \text{if } x \in W \\
                    0 & \text{if } x \notin W
                \end{cases}
            \]
            such that $\epsilon(g) = h$. Finally, we need to show that $\operatorname{im}(\iota) = \ker(\epsilon)$. If $f \in U$, then $\epsilon(\iota(f)) = 0$ since $\iota(f)$ is zero outside of $U$. Conversely, if $g \in V$ and $\epsilon(g) = 0$, then $g(x) = 0$ for all $x \in W$, which means that $g$ must be in the image of $\iota$. Therefore, we have shown that the sequence
            \[
                0 \xrightarrow{} U \xrightarrow{\iota} V \xrightarrow{\epsilon} W \xrightarrow{} 0
            \]
            is a short exact sequence.
        \end{proof}
    \end{enumerate}
    \item[] Remark(s): Since the problem stated that $U = \mathbb{F}(U)$, $V = \mathbb{F}(V)$ and $W = \mathbb{F}(Y \backslash X)$, I assumed that $\iota$ and $\epsilon$ were defined in terms of functions instead of elements of the vector spaces. If this is not the case, please let me know and I can adjust the proof accordingly.
\end{enumerate}


\begin{problem}
    \textit{Dimension}. Denote by $\operatorname{Vect}_{\mathbb{F}}^{fd}$ the collection of finite-dimensional vector spaces over $\mathbb{F}$, with linear transformations between them.
    \begin{enumerate}[a)]
        \item State the fact about uniqueness and existence of unique dimenstion function
        \[
            \dim: \operatorname{Vect}_{\mathbb{F}}^{fd} \to \mathbb{N},
        \]
        that satisfies certain desired properties. \\
        \textbf{Def.} For $V$ finite dimensional, the integer $\dim(V)$ is called the \underline{dimension} of $V$.
        \item Show that $\dim(M_n(\mathbb{F})) = n^2$.
        \item Suppose $1 + 1 \neq 0$ in $\mathbb{F}$. Consider the spaces $U = A_n(\mathbb{F})$, $V = M_n(\mathbb{F})$, $W = S_n(\mathbb{F})$, of anti-symmetric matrices ($A^T = -A$), all matrices, and symmetric matrices (sastisfy $A^T = A$), respectively. 
        \begin{enumerate}[i)]
            \item Show that, they form in a natural why a s.e.s.
            \item Deduce that $\dim(A_n(\mathbb{F})) = \frac{n(n-1)}{2}$ and $\dim(S_n(\mathbb{F})) = \frac{n(n+1)}{2}$.
        \end{enumerate}
    \end{enumerate}
\end{problem}

\begin{enumerate}[a)]
    \item 
\end{enumerate}



\end{document}